---
title: "cladeInference_2"
author: "Iskander Said"
date: "12/4/2020"
output: pdf_document
---

# README:

This example shows how to infer clades using test data. The SubFamily_inference module performs all necessary functions to hierarchically cluster alleles based on how correlated their copy numbers are across individuals in your data set. The pipeline() function defined in the following chunk wraps the main function of SubFamily_inference.R and iterates it across the 41 active TEs in our test data set. It includes vectors of the correlation cutoffs (1-r) and a boolean mask telling the module which TEs to excldue outliers for. To use this module the user must change the D_v, bool_mask vectors to match their dataset as well as modify any directories and file names.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("SubFamily_inference.R")

D_v <- c(0.5681367868025526,0.46425191203547655,0.4411858547936054,0.3362603560047991,0.4690155981177684,0.4279731548675725,0.47954672375772367,0.3828725421957896,0.493537112004956,0.4079762709005661,0.46283619082924043,0.10953532870101146,0.05877489009985082,0.25041520839980236,0.22458378676520008,0.11214905821852161,0.1612952508109864,0.4473320219906931,0.45279129094375314,0.44604417371355054,0.4197051313816731,0.2793291335504422,0.4423797589884111,0.5604722705971066,0.5286508819650785,0.398416314339164,0.30157738115657606,0.4802006073845314,0.292457988290345,0.19825250859657662,0.28026597172753276,0.2671602311703489,0.3311246702909195,0.5589724996606225,0.4305439804110055,0.4684870714100676,0.2812052156911843,0.4015464749075728,0.3974947619382654,0.34603367074832647,0.3289465851987603) #correlation cut-offs for hierarchical clustering as defined by dissimilarity distance (1-r)


pipeline<-function(clade_call, corr_cutoff){
  
  bool_mask <- c(FALSE,TRUE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE) #This array contains the mask of whether or not to remove copy number outliers from inference
  
  output=paste0('TEST_DATA/', haplotype_call) #change to your chosen output directory
  
  #create new directory
  if (dir.exists(output) ){
    
  }
  else{
    dir.create(output)
  }
  te_df <- read.csv('TEST_DATA/ACTIVE_TES_full.tsv', header = FALSE) #TE file data sheet containing the TE designations
  for (t in 1:length(te_df$V1)){
    
    input_path = paste0('TEST_DATA/CN_tables/FULL/') #path to your copy number CSVs
    outlier_path = paste0('TEST_DATA/CN_tables/NO_OUTLIERS/')
    TE=file.path(input_path,te_df$V1[t])
    out_mask = bool_mask[t]
    
    #Parameters:
    #TE_name -> full path leading to the allele copy number CSV
    #outlier -> T or F telling module whether or not to use an outlier excluded table
    #outlier_path -> path to CSV that has copy number outliers removed
    #output_dir -> output directory name
    #distance_cutoff -> 1-r; where to cut hierarchical clustering dendrogram
    #linkage -> clustering linkage method
    #color_pal -> color palette for visualization; length of vector MUST MATCH number of populations
    #alleleCN_dir -> directory of the copy number numpy files
    #sample -> population sample sheet
    
    extractHaplotypes(TE_name = TE, outlier = out_mask, output_dir = output, outlier_dir = outlier_path, linkage= 'average', dist_cutoff = corr_cutoff[t], color_pal=c('#6F0000','#009191','#6DB6FF','orange','#490092'),
                              alleleCN_dir="TEST_DATA/GDL_RUN-11-15-19", sample="TEST_DATA/GDL_sample_sheet.csv")
    
  }
}

```

Call the pipeline() wrapper function to infer clades from all 41 TEs in our test data:
```{r}
pipeline(clade_call = "CLADE_OUTPUT", corr_cutoff = D_v)
```

The wrapper will now run and produce graphical and tabular output in the desired directory.
